opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-05-14 chapters/01_introduction.tex
Added machine learning for spacegroup prediction papers copied from Henrik paper

Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement. \cite{Surdu2023,Feng2019} 
So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017} \cite{Lee2023} or, most recently, from generated synthetic crystals. \cite{Schopmans2023}

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
lattice parameters, spacegroup, and crystallite size from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. This includes noise, background, preferred orientation of the sample, ...
Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.

@@

Furthermore, even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent actual experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.
Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine learning approaches.
This is similar to the concept of style transfer\cite{Gatys2016, Ganin2015} found in the domain of machine learning applied to images.
Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.

>>>>

Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement. \cite{Surdu2023,Feng2019} 
So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals. \cite{Schopmans2023}

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
lattice parameters, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. This includes noise, background, preferred orientation of the sample, ...
Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.

@@

Furthermore, even completely unlabeled data can enable machine learning researchers to evaluate how closely their simulations represent experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.
Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine learning approaches.
This is similar to the concept of style transfer\cite{Gatys2016, Ganin2015} found in the domain of machine learning applied to images.
Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.

### 2024-05-14 chapters/03_SOTA_ML.tex
Added machine learning for spacegroup prediction papers copied from Henrik paper

>>>>

Using simulated diffractograms based on structures from the ICSD, Lee et al. managed to train a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool. \cite{Lee2020}
Park et al. reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms. \cite{Park2017}
A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{chitturi_automated_2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.

### 2024-05-14 main.tex
Commited changes by Henrik
Added todos in introduction
Commited updates from Henrik/Pascal
Pushed resolved comments
Added machine learning for spacegroup prediction papers copied from Henrik paper

\interfootnotelinepenalty=10000
% Changed idea title idea from orignal title of
% "Towards a free high quality powder x-ray diffractogram database"
% Name is supposed to be in analogy with Crystallography Open database
% which is the largest freely available database of crystal structures
\title{Experimental Powder X-ray Diffraction Open Database}

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[2]{Ben Breitung}
\author[2]{Simon Schweidler}
\author[3]{Andrea Hodge}
\author[3]{Adie Alwen}
\author[4,5]{Carolin Sutter-Fella}
\author[4]{Mriganka Singh}
\author[4]{Tim Kodalle}
\author[4]{Maged Abdelsamie}
\author[1,2]{Pascal Friederich}

@@

\maketitle
\begin{abstract}
\input{paper/00_abstract}
\end{abstract}

\section{Introduction}\label{sec:introduction}
\input{paper/01_introduction}
Test\cite{abadiTensorFlowLargeScaleMachine2015}

% Existing datasets (and why they dont cut it)
\section{Existing datasets}\label{sec:existing_datasets}

%\input{02_existing_datasets}
% Our dataset (and why its better)
\section{Our dataset}\label{sec:our_dataset}
%\input{03_our_dataset}

>>>>

\interfootnotelinepenalty=10000
% Changed idea title idea from orignal title of
% "Towards a free high quality powder x-ray diffractogram database"
% Name is supposed to be in analogy with Crystallography Open database
% which is the largest freely available database of crystal structures
\title{opXRD: Open Experimental Powder X-ray Diffraction Database}

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[1,2]{Jona Östreicher}
\author[2]{Simon Schweidler}
\author[3]{Adie Alwen}
\author[4]{Mriganka Singh}
\author[4]{Tim Kodalle}
\author[2]{Ben Breitung}
\author[4]{Maged Abdelsamie}
\author[4,5]{Carolin Sutter-Fella}
\author[3]{Andrea Hodge}
\author[1,2]{Pascal Friederich}

@@

\maketitle
\begin{abstract}
\input{chapters/00_abstract}
\end{abstract}

\newpage
\section{Introduction}\label{sec:Introduction}
\input{chapters/01_introduction}

\input{chapters/02_existing_datasets}

\input{chapters/03_SOTA_ML}

\section{opXRD database}\label{sec:our_dataset}
\input{chapters/04_our_dataset}

\section{How to contribute and how to use}\label{sec:how_to_use}
\input{chapters/05_how_to_use}

\subsection*{Data availability}
Data is available on xxx.

\subsection*{Conflicts of interest}
There are no conflicts of interest to declare.

\subsection*{Acknowledgements}
We acknowledge financial support by the German Research Foundation (DFG) through the Research Training Group 2450 “Tailored Scale-Bridging Approaches to Computational Nanoscience”. We acknowledge support by the Federal Ministry of Education and Research (BMBF) under Grant No. 01DM21001B (German-Canadian Materials Acceleration Center). We acknowledge financial support from the Helmholtz Foundation Model Initiative within Project "SOL-AI".
